---
title: "R_Markdown"
output: html_document
date: "2024-02-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set()
```

## 2.5 Markdown

# Title

## Main section

### Sub-section

#### Sub-sub section

Hyperlinks can be made like this: [QBIOL Masters](https://future-students.uq.edu.au/study/programs/master-quantitative-biology-5716?year=2021)

Latex: $$\frac{dN_{i}}{dt} = N_{i}(\frac{\mu _{max_{i}}R}{k_{i} + R} - \delta_{i} - d)$$

## 2.6 Code chunks

```{r}
norm = rnorm(100, mean=0, sd =1)
```

### Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(norm)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
```

Assigns a name to all generated plots and can also help with debugging.

```{r marsupial plot}
library(boilrdata)
plot(Mass ~ MaxLifeSpan, data=marsupials)
```

## 2.9 Tables in R markdown

```{r marsupial table}
library(knitr)
kable(head(marsupials))
```
